|

### 2021-07-27 backup1.Rmd

\section{Notation}
Let us fix the basic notation that will be employed henceforth. Upper case letters, e.g. $X_t, Y_t,$ denote random variables, while lower case letters 
denote the realizations. Finite sequences are made compact using the semi-colon notation, e.g. $x_{0:t}=(x_0, x_1, \dots, x_t)$ for some $t \geq 0.$ The sets of outcomes are denoted by upper case calligraphic letters, e.g. $\mathcal{X}, \mathcal{Y}.$  The $\sigma$-algebra
of a set $\mathcal{X}$ reads $\mathcal{B}(\mathcal{X}).$ The pair $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ is a measurable space.
The densities are represented with lower case cursive letters. For example, the density of $X_t$ in $x_t$ is $p_t^\theta(x_t),$ where $\theta\in\Theta$ is some (eventually known) vector, while the density 
of $X_t$ in $x_t,$ given that $X_{t-1}$ has previously taken value $x_{t-1},$ is $p_t^\theta(x_t\mid x_{t-1}).$  Note that $p_t^\theta(x_t\mid x_{t-1})$ is a member of the parametric family with parameters $(\theta, x_{t-1})$.
The probability distribution of a random variable $X_t$ is represented by $\mathbb{P}_t(dx_t),$ while the probability distribution of a sequence $X_{0:t}$ is represented by $\mathbb{P}_t(dx_{0:t}).$ The Dirac measure, i.e., the measure that assigns probability $1$ to the singleton $\{x\},$ is denoted by $\delta_x(dy)$.The expectation operator is denoted by $\mathbb{E},$ where the section of $\mathbb{E}$ at $\mathbb{Q},$ i.e., $\mathbb{E}_{\mathbb{Q}}$ denotes the operator computed with respect to the probability distribution $\mathbb{Q}.$ 
Let $\varphi$ be a function defined on the random variable $X\propto\mathbb{Q}.$ Then, we denote the expected value of $\varphi(X)$ with 
$$
    \mathbb{E}_{\mathbb{Q}}[\varphi(X)]=
    \int_{\mathcal{X}}\varphi(x)\mathbb{Q}(dx).
$$
Convergence in distribution is denoted with the symbol $\implies.$ So, $X\implies\N(0,1)$ means that the random variable $X$ converges in distribution to a normal Gaussian.

>>>>

\section{Notation}
Let us fix the basic notation that will be employed henceforth. Upper case letters, e.g. $X_t, Y_t,$ denote random variables, while lower case letters 
denote the realizations. Finite sequences are made compact using the semi-colon notation, e.g. $x_{0:t}=(x_0, x_1, \dots, x_t)$ for some $t \geq 0.$ The sets of outcomes are denoted by upper case calligraphic letters, e.g. $\mathcal{X}, \mathcal{Y}.$  The $\sigma$-algebra
of a set $\mathcal{X}$ reads $\mathcal{B}(\mathcal{X}).$ The pair $(\mathcal{X}, \mathcal{B}(\mathcal{X}))$ is a measurable space.
The densities are represented with lower case cursive letters. For example, the density of $X_t$ in $x_t$ is $p_t^\theta(x_t),$ where $\theta\in\Theta$ is some (eventually known) vector, while the density 
of $X_t$ in $x_t,$ given that $X_{t-1}$ has previously taken value $x_{t-1},$ is $p_t^\theta(x_t\mid x_{t-1}).$  Note that $p_t^\theta(x_t\mid x_{t-1})$ is a member of the parametric family with parameters $(\theta, x_{t-1})$.
The probability distribution of a random variable $X_t$ is represented by $\mathbb{P}_t(dx_t),$ while the probability distribution of a sequence $X_{0:t}$ is represented by $\mathbb{P}_t(dx_{0:t}).$ The Dirac measure, i.e., the measure that assigns probability $1$ to the singleton $\{x\},$ is denoted by $\delta_x(dy)$.The expectation operator is denoted by $\mathbb{E},$ where the section of $\mathbb{E}$ at $\mathbb{Q},$ i.e., $\mathbb{E}_{\mathbb{Q}}$ denotes the operator computed with respect to the probability distribution $\mathbb{Q}.$ 
Let $\varphi$ be a function defined on the random variable $X\thicksim\mathbb{Q}.$ Then, we denote the expected value of $\varphi(X)$ with 
$$
    \mathbb{E}_{\mathbb{Q}}[\varphi(X)]=
    \int_{\mathcal{X}}\varphi(x)\mathbb{Q}(dx).
$$
Convergence in distribution is denoted with the symbol $\implies.$ So, $X\implies\N(0,1)$ means that the random variable $X$ converges in distribution to a normal Gaussian.

### 2021-07-27 prova_knit_finale.Rmd

\section{State-Space Models}
We now define \textit{State-Space models} as Markov processes that are partially observed. Let $X_t\in\X$ and $Y_t\in\Y$ and let $\{(X_t,Y_t)\}$ be a stochastic process. To help intuition, suppose that $\{Y_t\}$ is observable, i.e., coincides with a sequence of observations, while $\{X_t\}$ is latent. Consider the initial distribution $\P_0(dx_0)$ and the probability kernels
$P_t(x_{t-1},dx_t)$ and $F_t(x_t,d_t)$, where $t=1,2,\dots$.
\
\begin{definition}\label{state-space-def}
The process $\{(X_t,Y_t)\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \P_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\P_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \P_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t).
\end{align*}
\end{definition}
From the second inequality, it is apparent that $\{(X_t,Y_t)\}$ is a Markov process with initial distribution $[\P_0(dx_0)F_0(x_0,y_0)]$ and transition kernel $P_t(x_{t-1},dx_t)F_t(x_t,dy_t).$ The third equality, instead, shows that, $\{X_t\}$ has a marginal distribution $$\P_T(dx_{0:T})=\P_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),$$ meaning that $\{X_t\}$ is marginally distributed as a Markov process with initial distribution $\P_T(dx_{0:T})$ and transition kernel $P_t(x_{t-1},dx_t).$

@@

  return(list(xs=xs,ws=ws,ess=ess))
}
```
We have already discussed the reasons why the SIS algorithm does not provide a good strategy in the filtering problem. We provide a graphical intuition of what happens when we use such filtering strategy on a simulated dataset. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. The results shown in the following two plots shows a clear degeneration of the effective sample size and bad fit of filtered states with respect to the true values.
```{r echo=FALSE}
set.seed(1991)
N=1000
sis<-SISfun(y,N,m0,C0,tau,sigma)
sisxs<-sis$xs
sisws<-sis$ws
sisess<-sis$ess
sis.data<-data.frame(df,sisxs,sisws,sisess)

@@

\begin{algorithm} BPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
These steps are resumed in our \code{PFfun} function.
\
\hrule
\hrule
\
\code{PFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\textbf{Arguments}

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
The estimated states of the Boostrap Particle Filter togheter with the effective sample size are shown in Figure \ref{fig:myfig3}. We decide to set $N=1000$,$m_{0}=0$,$C_{0}=100$ and $\tau=\sigma=1$. Notice how the resampling step allows the effective sample size not to drop, improving results.
```{r echo=FALSE, fig.align='center', fig.cap="Effective Sample Size", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
set.seed(1991)
N=1000

@@

\begin{algorithm} GPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item Compute $\sigma_{opt}^{2}$ 
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $\mu_{opt}$
\item Draw $x_{t}^{(i)} \sim N(\mu_{opt},\sigma_{opt}^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2+\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{GPFfun} function resume this passages.
\
\hrule
\hrule
\
\code{GPFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\textbf{Arguments}

@@

\begin{algorithm} APF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item For $k=1,...,N$:
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}f(y_{t}|g(x_{t-1}^{(i)})) $
\item Draw $x_{t}^{(k)} \sim N(x_{t-1}^{(I_{k})},\tau^2)$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|x_{t}^{(k)})}{f_{N}(y_{t}|g(x_{t-1}^{(I_{k})}))}$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{APFfun} function resume this passages.
\
\hrule
\hrule
\
\code{APFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\
\textbf{Arguments}

>>>>

\section{State-Space Models}\label{sec_state_space}
We now define \textit{State-Space models} as Markov processes that are partially observed. Let $X_t\in\X$ and $Y_t\in\Y$ and let $\{(X_t,Y_t)\}$ be a stochastic process. To help intuition, suppose that $\{Y_t\}$ is observable, i.e., coincides with a sequence of observations, while $\{X_t\}$ is latent. Consider the initial distribution $\P_0(dx_0)$ and the probability kernels
$P_t(x_{t-1},dx_t)$ and $F_t(x_t,d_t)$, where $t=1,2,\dots$.
\
\begin{definition}\label{state-space-def}
The process $\{(X_t,Y_t)\}$ is a State-Space model if the joint distribution of $(X_{0:T},Y_{0:T})$ is 
\begin{align*}
  \P_T(X_{0:T}\in dx_{0:T},Y_{0:T}\in dy_{0:T}) &= \P_0(dx_0)\prod\limits_{t=1}^T P_t(x_{t-1},dx_t)\prod\limits_{t=1}^T F_t(x_t,dy_t)\\
  &= [\P_0(dx_0)F_0(x_0,y_0)]\prod\limits_{t=1}^T[P_t(x_{t-1},dx_t)F_t(x_t,dy_t)] \\
  &= \P_T(dx_{0:T})\prod \limits_{t=1}^T F_t(x_t,dy_t).
\end{align*}
\end{definition}
From the second inequality, it is apparent that $\{(X_t,Y_t)\}$ is a Markov process with initial distribution $[\P_0(dx_0)F_0(x_0,y_0)]$ and transition kernel $P_t(x_{t-1},dx_t)F_t(x_t,dy_t).$ The third equality, instead, shows that, $\{X_t\}$ has a marginal distribution $$\P_T(dx_{0:T})=\P_0(dx_0)\prod_{t=1}^T P_t(x_{t-1},dx_t),$$ meaning that $\{X_t\}$ is marginally distributed as a Markov process with initial distribution $\P_T(dx_{0:T})$ and transition kernel $P_t(x_{t-1},dx_t).$

@@

  return(list(xs=xs,ws=ws,ess=ess))
}
```
We have already discussed the reasons why the SIS algorithm does not provide a good strategy in the filtering problem. We provide a graphical intuition of what happens when we use such filtering strategy on a simulated dataset. We decide to set $N=1000$, $m_{0}=0$, $C_{0}=100$ and $\tau=\sigma=1$. The following two plots show a clear degeneration of the effective sample size and bad fit of filtered states with respect to the true values.
```{r echo=FALSE}
set.seed(1991)
N=1000
sis<-SISfun(y,N,m0,C0,tau,sigma)
sisxs<-sis$xs
sisws<-sis$ws
sisess<-sis$ess
sis.data<-data.frame(df,sisxs,sisws,sisess)

@@

\begin{algorithm} BPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
These steps are resumed in our \code{PFfun} function.
\
\hrule
\hrule
\
\code{PFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\textbf{Arguments}

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
The estimated states of the Boostrap Particle Filter togheter with the effective sample size are shown in Figure \ref{fig:myfig3}. We decide to set $N=1000$, $m_{0}=0$, $C_{0}=100$ and $\tau=\sigma=1$. Notice how the resampling step allows the effective sample size not to drop, improving results.
```{r echo=FALSE, fig.align='center', fig.cap="Effective Sample Size", fig.height=4, fig.pos='H', fig.width=8, message=FALSE, warning=FALSE, results=F}
set.seed(1991)
N=1000

@@

\begin{algorithm} GPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item Compute $\sigma_{opt}^{2}$ 
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $\mu_{opt}$
\item Draw $x_{t}^{(i)} \sim N(\mu_{opt},\sigma_{opt}^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2+\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{1:t})$
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{GPFfun} function resume this passages.
\
\hrule
\hrule
\
\code{GPFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\textbf{Arguments}

@@

\begin{algorithm} APF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarize $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item For $k=1,...,N$:
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}=i) \propto w_{t-1}^{(i)}f(y_{t}|g(x_{t-1}^{(i)})) $
\item Draw $x_{t}^{(k)} \sim N(x_{t-1}^{(I_{k})},\tau^2)$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|x_{t}^{(k)})}{f_{N}(y_{t}|g(x_{t-1}^{(I_{k})}))}$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{1:t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{APFfun} function resume this passages.
\
\hrule
\hrule
\
\code{APFfun(data,N,m0,C0,tau,sigma,r)}
\
\hrule
\
\textbf{Arguments}
